---
title: "Exercise 12"
author: "Lili Benitez"
date: "2024-04-19"
output: html_document
---

```{r setup, echo=FALSE}
library(tidyverse)
library(naniar)
library(conflicted)
f<-"https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"
d<-read_tsv(f)
replace_with_na_all(d,   condition = ~.x == -999)
d<-d%>%
  dplyr::select(-refs, -`litter size`)

d_log<-d%>%
  mutate(across(where(is.numeric), log))

```

## Regression: 
Regress the (now log transformed) age [gestation(mo), weaning(mo), AFR(mo) (age at first reproduction), and max. life(mo) (maximum lifespan)] and mass [newborn(g) and wean mass(g)] variables on (now log transformed) overall body mass(g) and add the residuals to the dataframe as new variables (relGest, relWean, relAFR, relLife, relNewbornMass, and relWeaningMass).
```{r}

m_gest<-lm(data=d_log, `mass(g)`~`gestation(mo)`, na.action = na.exclude)# not padding NAs
relGest<-resid(m_gest)
m_w<-lm(data=d_log, `mass(g)`~`weaning(mo)`, na.action=na.exclude)
relWean<-resid(m_w)
m_AFR<-lm(data=d_log,  `mass(g)`~`AFR(mo)`, na.action=na.exclude)
relAFR<-resid(m_AFR)
m_life<-lm(data=d_log,  `mass(g)`~`max. life(mo)`, na.action=na.exclude)
relLife<-resid(m_life)
m_new<-lm(data=d_log,  `mass(g)`~`newborn(g)`, na.action=na.exclude)
relNewbornMass<-resid(m_new)
m_w_m<-lm(data=d_log,  `mass(g)`~`wean mass(g)`, na.action=na.exclude)
relWeaningMass<-resid(m_w_m)

d_log<-d_log%>%
  mutate(relGest=relGest, 
        relWean=relWean,
        relAFR=relAFR,
        relLife=relLife,
        relNewbornMass=relNewbornMass,
        relWeaningMass=relWeaningMass )

```

## Step 5
Plot residuals of max lifespan (relLife) in relation to Order. Which mammalian orders have the highest residual lifespan?
Plot residuals of newborn mass (relNewbornMass) in relation to Order. Which mammalian orders have the have highest residual newborn mass?
Plot residuals of weaning mass (relWeaningMass) in relation to Order. Which mammalian orders have the have highest relative weaning mass?
```{r}
ggplot(data=d_log, aes(x=order, y=relLife))+
  geom_boxplot()+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(data=d_log, aes(x=order, y=relNewbornMass))+
  geom_boxplot()+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(data=d_log, aes(x=order, y=relWeaningMass))+
  geom_boxplot()+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
Cetacea, Atiodactyla, Sirenia, and Perissodactyla, Proboscidea, have the highest residual life spans. Carnivora, Lagomorpha, Perissodactyla, Proboscidea, Tubulindentata have the highest residual newborn mass, although the latter seem to be data deficient. Sirenia, Tubulidentat, Xenarthra, Hyracooidea(these four data defic), Pholidota,Lagomorpha, Carnivora (very variable), have the highest relative weaning mass. 
## Step 6
Run models and a model selection process to evaluate what (now log transformed) variables best predict each of the two response variables, max. life(mo) and AFR(mo), from the set of the following predictors: gestation(mo), newborn(g), weaning(mo), wean mass(g), litters/year, and overall body mass(g).
HINT: Before running models, winnow your dataset to drop rows that are missing the respective response variable or any of the predictors, e.g., by using drop_na().

For each of the two response variables, indicate what is the best model overall based on AICc and how many models have a delta AICc of 4 or less?
What variables, if any, appear in all of this set of “top” models?
Calculate and plot the model-averaged coefficients and their CIs across this set of top models.

```{r}

library(MASS)
library(AICcmodavg)
library(MuMIn)
D_log<-d_log%>%
  drop_na(`gestation(mo)`,`max. life(mo)`, `AFR(mo)`, `newborn(g)`, `weaning(mo)`, `wean mass(g)`, `litters/year`, `mass(g)`)

####life span model selection####
life_1<-lm(data=D_log, `max. life(mo)`~ `gestation(mo)`+ `newborn(g)`+`weaning(mo)`+`wean mass(g)`+`litters/year`+ `mass(g)`,  na.action=na.fail)
summary(life_1) #OHHH lol the model wasn't including total mass because I had a comma, newborn and wean mass not sig
(s <- stepAIC(life_1, scope = . ~ ., direction = "both")) # - wean mass and - newborn(g) have better AIC
summary(s) 
life_2<-lm(data=D_log, `max. life(mo)`~ `gestation(mo)`+ `newborn(g)`+`weaning(mo)`+`litters/year`+ `mass(g)`,  na.action=na.fail)
summary(life_2)# newborn not sig
(s2 <- stepAIC(life_2, scope = . ~ ., direction = "both")) #  - newborn(g) has better AIC
life_3<-lm(data=D_log, `max. life(mo)`~ `gestation(mo)`+`weaning(mo)`+`litters/year`+ `mass(g)`,  na.action=na.fail)
summary(life_3) ##all sig

####Life span Model Averaging####
mods <- dredge(life_1, beta = "none", m.lim = c(0, 4))
(mods.list <- get.models(mods, subset = TRUE))
coef(mods)
(mods.avg <- summary(model.avg(mods, subset = delta < 4, fit = TRUE))) #two of the models have delta AICc below 4
confint(mods.avg)
(mods.avg <- summary(model.avg(mods, subset = cumsum(weight) <= 0.95, fit = TRUE)))

plot(mods.avg , full = TRUE, intercept = FALSE)

####AFR model selection####
AFR_1<-lm(data=D_log, `AFR(mo)`~ `gestation(mo)`+ `newborn(g)`+`weaning(mo)`+`wean mass(g)`+`litters/year`+`mass(g)`, na.action=na.fail)
summary(AFR_1) #newborn and wean mass and mass (g) are not sig
(step1 <- stepAIC(AFR_1, scope = . ~ ., direction = "both")) #dropping newborn, wean mass, or mass(g) all have highest AIC
AFR_2<-lm(data=D_log, `AFR(mo)`~ `gestation(mo)`+`weaning(mo)`+`wean mass(g)`+`litters/year`+`mass(g)`) #lets drop newborn (g) first
summary(AFR_2) #wean mass and mass (g) not sig, wean mass has highest p value
(step2 <- stepAIC(AFR_2, scope = . ~ ., direction = "both"))#dropping wean mass and mass g have best AIC
AFR_3<-lm(data=D_log, `AFR(mo)`~ `gestation(mo)`+`weaning(mo)`+`litters/year`+`mass(g)`) #lets drop wean mass
summary(AFR_3) #mass (g) not sig
(step3 <- stepAIC(AFR_3, scope = . ~ ., direction = "both"))
summary(step3) #keeping mass(g) is slightly better than dropping it, but AIC is within 2

####AFR Model averaging####
afr_mods <- dredge(AFR_1, beta = "none", m.lim = c(0, 4))
(afr_mods.list <- get.models(afr_mods, subset = TRUE))
(afr_mods.avg <- summary(model.avg(afr_mods, subset = delta < 4, fit = TRUE))) #7 models with delta AICc less than 4, model number 1236 is best but I can't go back and find it for some reason? 
confint(afr_mods.avg)
(afr_mods.avg <- summary(model.avg(mods, subset = cumsum(weight) <= 0.95, fit = TRUE)))
plot(afr_mods.avg , full = TRUE, intercept = FALSE)
```



